Vulnerability Analysis of Container Images in Neuroimaging

### 2020-06-23 paper.tex
Rewrote introduction

\begin{keywords}
Containers; Docker; Singularity; Security; Vulnerabilities; Image Scanning; Outdated packages; Shrinking Images; Neuroimaging.
\end{keywords}

@@

Containers, which provide an operating system (OS) level virtualization, are
popular. This is due to them being lightweight, flexible, and resource-efficient.
Additionally, they are capable of providing fast development cycles and continuous delivery with the
reduction of infrastructure cost~\cite{martin2018docker}. Containers are heavily utilised in various scientific fields,
in particular, to run big data analysis. OS-level virtualization gained a lot of interest when Docker~\cite{docker19} was
introduced in 2013. However, Docker is not supported by High Performance Computing (HPC)
environments, mainly due to the fact that Docker user can act as superuser inside the container.
As a result, Singularity~\cite{singularity19} was developed
for scientific applications by keeping the focus on HPC environments. In addition, Singularity also solves some other security issues
related to the architecture of Docker. It is used in many clusters including Compute Canada.
According to~\cite{kurtzer2017singularity}, Singularity is used in 36 research clusters.

\begin{comment}
\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{Figures/cluster_list.png}
	\caption{List of HPC clusters using Singularity~\cite{kurtzer2017singularity}.}
	\label{fig:cluster_list}
\end{figure}
\end{comment}

Inspite of their popularity, containers tightly
integrate with the host due to the sharing of the kernel, 
mounted file systems, and devices, which raises security
concerns.
In fact, in relation to other virtualization techniques, they share three main security threats~\cite{gantikow2016providing}:

\begin{comment}
\begin{figure}
  \centering
  \includegraphics[width=.7\columnwidth]{Figures/container.png}
  \caption{Architecture of Container-based
                Virtualization.}
  \label{fig:container-overview}
\end{figure}
\end{comment}
\textit{Privilege Escalation} A malicious attacker can break out of the container
and gain control of the host system and other containers running on the same host;

\textit{Denial-of-Service} One container may end up consuming all the resources of the
host, hence resulting in the starvation of the host and other containers;

\textit{Information Leak} Private data of the host and other containers could be
leaked and can be used for further attacks.

In order to mitigate these threats, containers use some features of the Linux kernel
such as \textit{control groups, namespaces}, etc. For instance, \textit{control groups}
are used to restrict the amount of resources that a container can use. Nonetheless, container's
security is still one of the concerning issues.
%However, two features of the Linux kernel, \textit{control groups} and \textit{namespaces},
%are used by containers to mitigate these threats to a certain extent. The security
%of containers is still the most concerning issue. 
According to a 2015 Forrester survey, security is the
biggest concern when deploying containers~\cite{bettini2015vulnerability}.
%Containers are popular in various scientific fields, in particular to run Big Data analyses
%on HPC clusters. 
%However, deploying Docker container images on clusters raises security issues. Some security issues 
%that could possibly lead to privilege escalation, DoS or information leak. Some
%Above mentioned threats are not the only security issues that containers may suffer.
%Other security issues are related to the architecture of Docker but have been 
%largely solved by Singularity.
%In fact, Singularity was developed for scientific applications
%by keeping the focus on HPC environments. Although, this work is of general interest,
%our particular focus in this paper is neuroimaging as containers are popularly used for scientific computing and data science
%in neuroscience.
While containers still suffer general security problems that may affect any application using
them, our particular focus in this research is the security of container images.
In this research work, we aim to study and analyse the security of container images.
Although this work, is of general interest and can be applied on any field, we focus in this
paper on neuroimaging. This is due to the fact that containers are particulary popular in
scientific computing and heavily utiilised in neuroscience, which makes it a good candidate
for this research.
This popularity of containers in neuroscience leads us to the following
primary research question:

\textit{RQ1: What is the current security state of
neuroscience container images that are deployed on HPC clusters?}

In fact, answering this question leads us to further questions as described below. 
To answer our primary research question, we scan neuroscience images. The purpose
of this scanning process is to collect information about the container images
and then investigate these findings against security vulnerabilities.
%we scan images that are used on Compute
%Canada clusters by the Neuroscience field. Scanning refers to the practice of
%collecting all information about the container image, investigating it
%for vulnerabilities, and finally producing a
%report to summarize scanning results. 
It is broadly categorized into
two types: \textit{Static} and \textit{Dynamic} scanning. The former
involves extracting information of an image without running that image into a container,
whereas the latter involves running a container with the target image.
While scanning may provide different information, here, we focus mainly on security vulnerability.
These types of vulnerabilities may be caused by an exploitable weakness in developer's software.
Such exploitation may result in serious attacks such as gaining unauthorized access on the target
computer, deletion of mounted files, etc. In this work, we target the detection of vulnerabilities
that are present in neuroscience container images. It is further important to answer following three 
secondary questions:
%Vulnerability is a weakness in the developer's software that can be exploited
%by the attacker to perform unauthorized actions on the target computer.
%In this paper, we report vulnerabilities that are present in
%container images, specifically used by neuroimaging field. Particularly,
%we aim to answer four primary research questions:

\textit{RQ2:} Can vulnerabilities be reduced by updating the images?

\textit{RQ3:} Can vulnerabilities be reduced by shrinking the size of the images?

\textit{RQ4:} What is the effect of both updating
and shrinking image size on vulnerabilities?

%%%%%%%%%Related Work%%%%%%%
%The existing research work on containers focuses mainly on the security of Docker
%containers.
%This focus is justified by the fact that containers expose the host's resources
%(e.g., file system/ IPC) to the guest system. This feature raises a confidentiality
%threat for the applications running on the same host. Previous studies evaluated
%the security of Docker engine ~\cite{martin2018docker, sultan2019container, combe2016docker, bui2015analysis},
%and have scanned vulnerabilities on Docker hub images~\cite{Shu2017, gummaraju2015over}.
%Zerouali et al. in~\cite{zerouali2019relation} studied severity vulnerabilities and bugs in Docker 
%containers due to outdated packages. 
%
%There are various image scanners available for scanning Docker images, for
%instance Anchore, CoreOS's Clair, etc.
%However, these scanners cannot be used to scan Singularity images due to different image formats.
%Therefore, Singularity came up with Stools~\cite{stools}, which can be used for scanning Singularity images
%for security and quality checks.

This paper is organized as follows: in the next section, we provide a description
of materials and tools, which we utilise throughout our experiments. We also present
different image scanners used for vulnerability reporting in container images.
The following section of this paper provides our approach and methodology, where we
mainly examine the effect of image update and image reduction for the purpose
of reducing vulnerabilities. Afterwards, we provide the results of our experiments,
where we indicate a number of detected vulnerabilities in container images used for
data analysis on shared HPC clusters. The last two sections provide a conclusion of this
work following by a summary and an indication of our future work.

@@

>>>>

\begin{abstract}
	\tristan{Abstract goes here}
\end{abstract}

\begin{keywords}
Containers; Docker; Singularity; Security Vulnerabilities; Neuroimaging.
\end{keywords}

@@

% Containers are popular on HPC, in particular Singularity
Software containers have emerged has an efficient solution to deploy
scientific data analyses on High Performance Computing (HPC) clusters, due
to their portability, ease of use, and limited overhead. The
Singularity~\cite{kurtzer2017singularity} framework is often preferred to
Docker\footnote{\url{http://docker.com}} on HPC clusters, due to its
secure handling of multi-user environments, and its convenient support for
Docker images. Singularity is now available in dozens of HPC
clusters around the globe, and used routinely for Big Data analyses.

% Containers are meant to isolate applications from the host, but security remains an issue
Using core features of the Linux kernel such as namespaces, control groups
and chroot, containers isolate processes from the
host computer,  
and control the memory, CPU, network and file-system resources assigned to
them. However, containers still share the kernel, mounted file systems and
some devices with the host, which raises potential security issues. In
particular, privilege escalation, denial of service and information leak
are always plausible~\cite{gantikow2016providing}. Therefore, security remains 
an important concern when deploying containers~\cite{bettini2015vulnerability}.

% Container images, our study
This study is a vulnerability analysis of container images used in HPC
clusters. Container images contain complete operating-system distributions,
including the data analysis software and its required dependencies.
Images are rarely updated, as they are often built once
for the entire duration of an analysis, to avoid results perturbations
coming from software updates~\cite{gronenschild, glatard2015}. Images also
commonly include more dependencies than required by a specific analysis, to 
enable reusability across experiments.

Our research questions are the following:

\textit{RQ1: What is the current amount of known vulnerabilities in
container images deployed on HPC clusters?} Vulnerabilities are possible
attack vectors that may seriously compromise the security of HPC clusters
and the integrity of user data. We report vulnerability analyses produced
by four popular image scanning tools, Anchore, Vuls, Clair and Stools.

\textit{RQ2: Can the amount of vulnerabilities be reduced by updating the images?}  
For many reasons related to reproducibility and the life-cycle of research
projects, container images deployed on HPC clusters often include outdated
software. We report on the effect of software updates on the vulnerability
analysis of RQ1.

\textit{RQ3: Can the amount of vulnerabilities be reduced by shrinking the images?} 
Container images often include more software packages than required for a
typical analysis. We report on the impact of this additional software on
the presence of vulnerabilities.

We answer these questions taking the particular example of neuroimaging, a
domain that has entered the Big Data era as many other
ones~\cite{vanhorn-toga}. We focus on the analysis of container images
present in BIDS apps~\cite{xxx} and Boutiques~\cite{xxx}, two containerized
application frameworks commonly used in neuroimaging.

The next section describes the materials and tools used in the experiments,
including image scanners and container images. We then describe our image
update and shrinking methodology to reduce vulnerabilities. Results review
the detected vulnerabilities in container images, explain observed
discrepancies between scanners, and report on the efficiency of image
update and size reduction. As a conclusion, we provide a set of image
creation guidelines for a secure deployment on HPC clusters.

@@

\section{Discussion}

% Would the results apply to other domains than neuroimaging
